---
title: "STAT 27850 Project 1"
header-includes:
  - \usepackage{soul}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \DeclareMathOperator*{\argmin}{argmin}
output:
  pdf_document: default
geometry:
- margin=1 in
- top=  0.9cm
fontsize: 10pt 
---
```{r global-options, include=FALSE}
knitr::opts_chunk$set(
        fig.width=5, fig.height=3.4,
        fig.path='Figs/', fig.align='center',
        warning=FALSE, message=FALSE,
        strip.white=TRUE, cache=TRUE)
```
\newcommand{\h}[1]{\colorbox{yellow}{$\displaystyle #1$}} 
\titlespacing{\title}{0pt}{\parskip}{-\parskip}    
\vspace{-12.5truemm} 
 
\begin{center}
Christopher Tang
\end{center}


**1**
```{r, eval = TRUE}
library(dplyr)
library(ggplot2)

load("bikedata.RData")
# format covariates into a single data frame for easy viewing
bikenum <- as.data.frame(bikenum); day_of_week <- as.data.frame(day_of_week)
duration <- as.data.frame(duration); member <- as.data.frame(member)
station_end <- as.data.frame(station_end); station_start <- as.data.frame(station_start)
days_since_Jan1_2010 <- as.data.frame(days_since_Jan1_2010)
data <- cbind(station_start,station_end,duration,member,bikenum,day_of_week,days_since_Jan1_2010 )
data$ridename <- paste(station_start,' - ',station_end)
data <- add_count(data,station_start,station_end)
```

```{r}
# add weather covariates
weather <- read.csv("WeatherData.csv")
data <- left_join(data,weather,by='days_since_Jan1_2010')
```

   \emph{(a)}     
       


```{r}
# top ride
df <- data[data$station_start==31104 & data$station_end==31106,] %>%
  group_by(days_since_Jan1_2010) %>%
  summarize(mean_daily_duration = mean(duration), count=n(), day_of_week=day_of_week, member=member)

ggplot(data=df, aes(x=days_since_Jan1_2010, y=mean_daily_duration)) +
  geom_line() + ylim(c(0,1000))
```
The average daily duration oscillates over time very drastically, suggesting people happen to ride longer bike trips on certain days of the week.


## Subset data and run regression on each ride and save p-vals + coefficients
```{r}
# routes without infrequent rides
routes <- data[data$n > 100,][,c('station_start', 'station_end')] %>% distinct()
route_pvals <- rep(NA, nrow(routes))
route_coef <- rep(NA, nrow(routes))

fit_list <- vector(mode = "list", length = nrow(routes))
for (r in 1:nrow(routes)) {
  # remove outlier rides > 3 sd away from route mean
  route_data <- data[data$station_start==routes[r,]$station_start & data$station_end==routes[r,]$station_end,]
  upper_cutoff <- mean(route_data$duration) + 3 * sd(route_data$duration)
  lower_cutoff <- mean(route_data$duration) - 3 * sd(route_data$duration)
  route_data <- route_data[route_data$duration < upper_cutoff & route_data$duration > lower_cutoff,]
  
  # run regression of route 
  fit <- lm(duration ~ days_since_Jan1_2010 + day_of_week + member + TMAX + PRCP, data=route_data)
  pval <- summary(fit)$coefficients[,4]['days_since_Jan1_2010']
  coef <- summary(fit)$coefficients[,1]['days_since_Jan1_2010']
  route_pvals[r] <- pval
  route_coef[r] <- coef
  fit_list[[r]] <- fit
}
```

```{r}
save('fit_list', 'route_pvals', 'route_coef', file="data/regression_out.RData")
```


```{r}
# routes w/ p val lower boferonni at alpha= 0.05
print(routes[which(route_pvals < 0.05/length(route_pvals)),])

print(route_coef[which(route_pvals < 0.05/length(route_pvals))])
# stations[stations[,1] == 31011]
# stations[stations[,1] == 31007]
# stations[stations[,1] == 31217]
# stations[stations[,1] == 31235]
# stations[stations[,1] == 31200]
# stations[stations[,1] == 31225]
```

```{r}
df <- data[data$station_start==31200 & data$station_end==31225,] %>%
  group_by(days_since_Jan1_2010) %>%
  summarize(mean_daily_duration = mean(duration), count=n(), day_of_week=day_of_week, member=member)

ggplot(data=df, aes(x=days_since_Jan1_2010, y=mean_daily_duration)) +
  geom_line() #+ ylim(c(0,5000))

# looks like outlier
```



       
